Type Error Telemetry at Scale

### 2023-09-22 programming23/paper.tex
related work: telemetry, privacy ... more coming

Liblit dissertation~\cite{liblit-thesis}

@@

Differential privacy for coverage analysis of software traces,
adds noise to traces,
hard because infinitely many traces,
use CFG abstraction of possible behaviors for callgraph analysis,
count sketch data structure~\cite{hlzbr-ecoop-2021}.

Nasko~\cite{zhlbr-cc-2020,zhlbr-oopsla-2020,hlzbr-ecoop-2021}

@@

\cite{bgimgm-cse-2016}
study Java compiler error messages after an intervention, counting num errors,
counting repeats, counting per user

\cite{anna-russo-kennedy-ms-2006}
BitFit, web-based java ide log num hint requests num, result of check answer
requests num compile, num run, no code?! good we have the same constraint

\cite{ab-sigcse-2015}
frequency, time-to-fix, spread of errors
 (we can't do time)
parsing compile error + source code
 sometimes with custom parsers

\cite{m-masters-2016}
study errors in blackbox dataset,
 overwhelmingly syntax

\cite{bkmu-sigcse-2014}
blackbox, 100k users
source code, diffs
error line, column, message;
keep all comments EXCEPT one above the class declaration

\cite{bask-icer-2018}
2TB data, Java programs
18 publications survey, technical challenges in analysis
 most pubs analyze errors;
nobody asked:
 which exns do students encounter (what?! mccall kolling 2014)
 execution / editing timelines
(blackbox analysis tutorial in sigcse 2020 on a mini dataset, to help researchers start writing an analysis)

\cite{t-hatra-2021}
plan for a large user-centered study 
... not very deep, look at edit sequences, replay for errors

@@

\cite{sscwj-oopsla-2017}
NATE numerical analysis of type errors
5000 labeled programs
 errors from class, first fix

\cite{h-dissertiation-2005}
constraint-based type inference;
general te criteria:
- target audience (novice / expert)
- output format (text / viz)
- interactivity (yes / no)
- heuristics (yes / no)
- primary-or-external (most type errors are easy, some could really use dedicated help)

\cite{hw-scp-2004}
location of type error as a slice,
 not a srcloc,
 not a subtree

\paragraph{Research on Errors}

TODO Liblit dissertiation

Mind your language~\cite{mfk-onward-2011}.

\paragraph{Telemetry}

\begin{table}[t]
  \caption{Comparing telemetry systems}
  \label{t:telemetry-design}

  \begin{tabular}{l@{~}cccccc}
    &             & PII       & Session ID & Deterministic & Private Data (???) \\\midrule
    & Transparent & \chkNo    & \chkNo     & \chkYes       & \chkNo      \\
  * & Us          & \chkNo    & \chkYes    & \chkNo        & \chkYes     \\
    & Diff.Priv   & $\epsilon$ & \chkYes   & \chkYes       & \chkYes     \\
    & Liblit      & ?         & ?          & ?             & ?           \\
    & \code{.NET} & \chkMaybe & \chkYes    & \chkYes       & \chkYes     \\
    & VS Code     & \chkYes   & \chkYes    & \chkYes       & \chkYes     \\
  \end{tabular}
\end{table}

Transparent telemetry: explain what it is and contrast our approach.
What essential, non-transparent things did we collect?

counting only, public decisions about what to count, public data,
collected weekly for a sample of clients, LastWeek field = prior day when
system gathered any data;
eg command invocations, lib stack frames, 
no user ID, no machine ID,
no time-ordered traces~\cite{transparent-telemetry}.

kindle track every tap (page turn etc),
enables whispersync feature,
drove design of navigation tools,
opt-out possible~\cite{kindle-telemetry}

vscode: usage data, crash reports, error data (not a crash but unexpected);
save file, open terminal, copy-paste, autocomplete offered, git queries, machine id,
session id, timestamp,
opt-out may be possible, but not for all usage data (per license, sec 2a) and
every extension can do its own thing~\cite{vscode-telemetry}

.NET SDK and .NET CLI:
crash reports, command invoked, args, hashed cwd, timings;
more listed online;
custom builds beware inadvertant disclosure, keep your filepaths clear;
data published in aggregate under CC-BY;
opt-out possible~\cite{dotnet-telemetry}

>>>>

\begin{table}[t]\centering
  \caption{Comparing telemetry systems}
  \label{t:telemetry-design}

  \FILL{}

  \begin{tabular}{l@{~}cccccc}
    &               & Code      & Session ID & Timestamp     & Deterministic (???) & Private Data (???) \\\midrule
    & Transparent   & \chkNo    & \chkNo     & \chkNo        & \chkYes       & \chkNo      \\
  * & Roblox Studio & \chkNo    & \chkYes    & \chkNo        & \chkNo        & \chkYes     \\
%% orthogonal %%    & Diff.Priv     & $\epsilon$ & \chkYes   & \chkYes       & \chkYes       & \chkYes     \\
%% orthogonal %%    & CBI           & \chkMaybe & \chkMaybe  & \chkNo        & \chkNo        & ?           \\
    & Classic       & \chkYes   & \chkYes    & \chkYes       & \chkYes       & \chkYes     \\
  \end{tabular}
\end{table}

% kindle track every tap (page turn etc),
% enables whispersync feature,
% drove design of navigation tools,
% opt-out possible~\cite{kindle-telemetry}

Type error telemetry in Roblox Studio has two distinguishing characteristics:
it includes no personal information (PII), and it sends data on
randomly-selected runs of the type checker rather than specific events of interest.
To the best of our knowledge, this design occupies a unique position relative to prior
work~(\cref{t:telemetry-design}).

In other professional-grade IDEs, such as VSCode~\cite{vscode-telemetry}
and \code{.NET}~\cite{dotnet-telemetry}, telemetry may include all sorts of
data: error reports, source code fragments, timestamps, and filepaths.
Users may be able to opt out of some telemetry, but the details depend on the
license agreement.
Furthermore, at least in VSCode, IDE extensions can report their own telemetry.
While this data can be invaluable for discovering bugs in production, it
must be handled with extreme care.

The BlackBox project similarly collects telemetry in broad strokes.
Every Java program submitted for compilation goes into the BlackBox database (minus
any comments above the main class) along with the compiler's response~\cite{bkmu-sigcse-2014}.
Though somewhat intrusive, the millions of programs in the BlackBox dataset
have enabled dozens of contributions to the area of CS
education~\cite{bask-icer-2018}, ranging from categorization of novice
errors~\cite{mk-fie-2014,m-masters-2016} to analyses of error frequency and
time-to-fix~\cite{ab-sigcse-2015}.

At the opposite end of the privacy spectrum, Go's design for \emph{transparent
telemetry} reports only counter values~\cite{transparent-telemetry}.
Unlike Roblox Studio, transparent telemetry includes no timestamps and no 
session ids.
The data is useful for learning about the frequency of specific events and the
distribution of events across clients (e.g., how many \code{CodeTooComplex}
errors arise? what is the average number of type errors?).
However, the lack of timestamps and ids means that it cannot track edits
over time (a feature that is quite useful to us, for example, in~\cref{f:error-density}).
% In~\cref{f:error-density}, for example, we see that increases to the number
% of type errors are quickly followed by decreases.
% Without timestamps, we would be unable to rule out the possibility that users
% build up a mountain of type errors and then slowly reduce them to zero.

One way to futher strengthen the privacy of any telemetry design, including
transparent telemetry, is to fuzz the data using techniques from
differential privacy~\cite{zhlbr-cc-2020,epk-ccs-2014,wblj-usenix-2017}.
Recent work on fuzzing traces~\cite{zhlbr-cc-2020},
accounting for known relationships between events~\cite{zhlbr-oopsla-2020},
and performing the fuzzing on the client side~\cite{hlzbr-ecoop-2021} is
especially relevant to the session-based data that Roblox Studio collects.
Applying differential privacy is subtle, however.
While we can likely fuzz the total number of type errors, if we were to
fuzz the number of \code{CodeTooComplex} errors carelessly, it might drastically
affect our conclusions.

Lastly, cooperative bug isolation is a related method for designing
telemetry systems~\cite{liblit-thesis}.
%% \FILL{} adaptive bug isolation~\cite{nl-icse-2010}, search on the control dependence graph.
%% 1\% overhead vs 87\% for sampling and 300\% for binary instrumentation
The focus is not on privacy, but rather on how to collect a small amount of
data from each user and perform statistically-sound analyses.
Each feature of interest \emph{within} a user's codebase has a uniform-random
probability of contributing telemetry.
The system-builder can analyze this telemetry with predicates to identify
notable events and logistic regression to narrow down nondeterministic bugs.
There is a high burden on designers to decide what to capture,
but a careful design can minimize the data from each individual user.
%% more in diss: how to deal with native compilers, threads, dynamic loading
%% ... privacy is future work, can maybe avoid private data with a type system

\paragraph{Telemetry Rates}

We chose randomly, uniform distribution.

\paragraph{TBD}

@@

@@

\paragraph{Contrast: User Studies and Types and/or Proofs}

\cite{h-dissertiation-2005}
heeren
constraint-based type inference;
general te criteria:
- target audience (novice / expert)
- output format (text / viz)
- interactivity (yes / no)
- heuristics (yes / no)
- primary-or-external (most type errors are easy, some could really use dedicated help)

\cite{sscwj-oopsla-2017}
learning to blame localizing type errors
NATE numerical analysis of type errors
5000 labeled programs
 errors from class, first fix

\cite{rsgl-cpp-2020}
user study coq proof assistant
N devs, XXX

@@

\cite{hw-scp-2004}
haack wells
location of type error as a slice,
 not a srcloc,
 not a subtree

\paragraph{Errors}

Mind your language~\cite{mfk-onward-2011}.
60 students (from 120, self-selected),
50 min per week for 6 weeks,
edit sequences,
followup interviews,
huge problem: vocabulary (function body, arity),
problem: semantics of highlight,

\cite{anna-russo-kennedy-ms-2006}
BitFit, web-based java ide log num hint requests num, result of check answer
requests num compile, num run, no code?! good we have the same constraint
115 bitfit users, 728 sessions

\cite{bgimgm-cse-2016}
user study,
study Java compiler error messages after an intervention, counting num errors,
counting repeats, counting per user
changing errors helps

\cite{dlc-citicse-2014}
enhanced stx error seems ineffectual

- - -

\cite{t-hatra-2021}
plan for a large user-centered study of proof assistants,
events: up, down, to cursor, interrupt, reset.
... not very deep, look at edit sequences, replay for errors

\cite{wk-koli-2020,lgk-pj-2022} 
even more edit sequences
b2t2, examplar, ... must be others outside brown!!
